---
title: "temp"
author: "Lisa OConnell"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---


REMEMBER TO CHANGE THE OUTPUT TYPE above AND CREATE RMD files

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Introduction  
Data from the American Business Survey (ABS) conducted by the US CENSUS Bureau is explored in this vignette. The purpose of the vignette is to demonstrate access to multiple endpoints within the ABS data using a __X__ query parameters.  Additionally an Exploratory Data Analysis (EDA) is completed which includes both numeric and graphical summarizations.  This vignette was created and is shared using a GitHub repository and GitHub page.  


# Description of Data  
__Endpoints Over Time__  
The American Business Survey data represents aggregated responses from the ABS survey conducted in the prescribed year from the selected endpoint.

The ABS data accessed for this vignette represents the endpoints for the year 2021 which contains data from the 2020 survey.  Additional endpoints are available for the years 2020-2018.   


__Endpoint Pathways:__  
-  Company Summaries: http://api.census.gov/data/2020/abscs   
-  Characteristics of Business: https://api.census.gov/data/2020/abscb  

  
*Additional Endpoints Available*  
An additional endpoint is available, but is not used in this vignette:  
-  Characteristics of Business Owners: https://api.census.gov/data/2020/abscbo  
- Module Business Characteristics: https://api.census.gov/data/2020/absmcb  

__Data Categorizations  
The ABS contains summarized data from survey respodants.  Data is available at the US, state and statistical area levels. This vignette uses data at the US level.  To provide a profile of businesses in America, responses are tabulated across several categories.  

For this vignette, I will be looking a the following categories for Company Summaries:
Industry, Sales/Receipts and reported Veteran Status

For the Business Owner data, I will look at......

Both datasets can be further profiled by using survey responses classifications related to the business owner including for example veteran stats, sex, & race.

Quantitative Variables for both endpoints include the number of employees and annual payroll expenses.  


__Data Format:__
As described in the ["CENSUS Data API User Guide"](https://www.census.gov/content/dam/Census/data/developers/api-user-guide/api-guide.pdf) The American Business Survey data returned through the API is a 2 dimensional JSON array formatted as shown below for 3 variables:  

[["EMP" ,"STATE", "PAYANN"]],  
["10000, "GA", "500000"],  
["5000, "TX", "650000"].... 

Due to the way the categorical information is stored, the most practical way to efficiently deal with incoming data is to execute calls within the AP for the slices of data of interest.  As a result, a general function is not used to query all classifications in one data pull.  The tabulations by category are appended to the industry records making a full file difficult to parse. 


# Requirements: R Libraries  
The following libraries are used to process access and process data in R.  The `tidyverse` library provides a set of packages relevant for data tidying, wrangling, and summarization.  The `httr` package provides functions to work with urls.  The `jsonlite`package provides functions to work with json data.  

```{r libraries}
library(tidyverse)
library(httr)
library(jsonlite)

```


# Functions to Retrieve ABS Data

Several functions are created to facilitate data retrieval. These functions allow query arguments to be passed in to return specific results.  A wrapper function called `GetData` is created to contain the subfunctions needed to execute the API call (`GetAPIPath`,`GetMyRawData`), convert the data from JSON (INSERT FX), create a tidy tibble (`CreateTibble`), and reformat variables (`ChangeVars`).  

The `GetAPIPath` function allows for multiple arguements to be passed in to specify query terms.  For the two endpoints used in this vignette, the following query parameters are allowed:

*Characterics of Business data:*  
Query parameters:

__NAICS2017__   The NAICS industry code from the 2017 NAICS table. 

__VET_GROUP__ A categorical variable that returns the status of the business owner as a veteran INSERT VALUES HERE

__SEX__ - A categorical variable returning the sex of the business owner.

__Survey Questions: Family Owned Businesses__ - A survey response value captured in the survey questions variables.  


Naming Conventions:  
-Camelcase for variables and Functions  

-SnakeCase for Datasets  


COMMENT FOR BLOG POST: The useful videos on CENUS website, and the ability to get more informative error codes by running the GET command directly into a browser.

These functions will create a URL path the endpoint that contains the user input arguments for queries.

Functions were built to have user friendly arguments which convert user input into ABS variable names and codes.  



## HELPER FUNCTIONS


```{r CSDataSat}
#USE THIS!

CSData<- function(vetStatus="*",industryCode="*") {
#defaults are wildcard values
#Defaults used to pull these variables when not specified by user. 
endpath <- "https://api.census.gov/data/2020/abscs?"

#A list of variable to return, excluding the query variables
varlist <- "get=FIRMPDEMP,PAYANN,EMP,SECTOR,SUBSECTOR"

#Parse the user-friendly 

#Paste query variables to path
path <- paste0(endpath,
              varlist,
            #A list of potential query arguments
              "&for=us:*",
              "&NAICS2017=",
               industryCode,
              "&VET_GROUP=",
               vetStatus, 
              "&key=324a30368aa24fca02eb737c15d7c32c63c891cb")
 
raw_data <- GET(path)
#Print status code of returned data
a<-status_code(raw_data)
print(a)

my_df<-raw_data %>%
     Wrangle() %>%
     #Modify variable types
     mutate(Firms = as.double(FIRMPDEMP), 
            Payroll = as.double(PAYANN),
            Employees = as.double(EMP)) %>% 
      mutate(Industry = as.factor(NAICS2017), 
             SECTOR=as.factor(SECTOR), 
             SUBSECTOR= as.factor(SUBSECTOR),
             VET_GROUP= as.factor(VET_GROUP)) %>% 
      DeAgto4()

return(my_df)

}
```


```{r FxConvertData}
ConvertData<-function (myraw) {
  a1<-fromJSON(rawToChar(myraw$content))
  a2<-as_tibble(a1)
  invisible(a2)
}

```

```{r FxCreateTibble}
CreateTibble<-function(temp_df) {
  #Create tibble 
  temp_df<-as_tibble(temp_df)

  #Get variable names from the first row of data
  var_names<-slice(temp_df,1)
  num_cols<-length(var_names)
  b<-pivot_longer(var_names, cols=1:num_cols,values_to = "new_name", names_to="old_name")
  #Replace variable names
  my_names<-b$new_name
  my_names

  #Remove the column name header row
  colnames(temp_df)<-my_names
  new_data<-temp_df[-1, ]
  #Return data
  invisible(new_data)
}
```

```{r WrangleWrapper}
Wrangle<- function(my_df){
  tidy_data <- my_df %>% 
            ConvertData() %>% 
            CreateTibble()
  invisible(tidy_data)
}

```

```{r FXChangeVarsCS}
#NOT USING SINCE CHANGED HOW DATA IS BEING PULLED

#Function specific to the variables in the Characteristics of Business dataset
#Converts all numeric variables from char to numeric
#Converts categorical variables to factors

ChangeVarsCS<-function(my_df) {
  #Create vector of numeric variables
  number_vars<-c("PAYANN", "FIRMPDEMP", 
               "EMP", "RCPPDEMP")        
  
  #Create vector of factor variables 
  factor_vars<-c("EMPSZFI", "RCPSZFI", 
                "SEX", "VET_GROUP", "ETH_GROUP", "RACE_GROUP")
  #Apply changes to my_df
  my_df<-my_df %>% 
      mutate(across(factor_vars,as.factor)) %>% 
      mutate(across(number_vars, as.double))
return(my_df)
}  
```

```{r FXChangeVarsCB}
#Function specific to the variables in the Characteristics of Business dataset
#Converts all numeric variables from char to numeric
#Converts categorical variables to factors

ChangeVarsCB<-function(my_df) {
  #Create vector of numeric variables
  number_vars<-c("PAYANN", "FIRMPDEMP", 
               "EMP", "RCPPDEMP")        
  
  #Create vector of factor variables 
  factor_vars<-c("EMPSZFI", "ETH_GROUP", 
                 "SEX", "VET_GROUP", "RACE_GROUP", "SECTOR","SUBSECTOR")
  #Apply changes to my_df
  my_df<-my_df %>% 
      mutate(across(factor_vars,as.factor)) %>% 
      mutate(across(number_vars, as.double))
invisible(my_df)
}  

```


```{r FXChangeVarsCBFamily}
#Function specific to the variables in the Characteristics of Business dataset
#Converts all numeric variables from char to numeric
#Converts categorical variables to factors

ChangeVarsCBFamily<-function(my_df) {
  #Create vector of numeric variables
  number_vars<-c("PAYANN", "FIRMPDEMP", 
               "EMP", "RCPPDEMP")        
  
  #Create vector of factor variables 
  factor_vars<-c("SECTOR","SUBSECTOR","SEX")
  #Apply changes to my_df
  my_df<-my_df %>% 
      mutate(across(factor_vars,as.factor)) %>% 
      mutate(across(number_vars, as.double))
invisible(my_df)
}  

```


```{r FXDeaggregate}
#Removes aggregated records to 4 digit NAICS codes.
#This is needed to remove duplicate counts that can occcur when the wildcard* is used.
DeAgto4<-function (a_df) {
    a <- a_df %>%
        filter(((nchar(NAICS2017)) > 3) ) %>% 
        #filter((substring(NAICS2017, 3,3)!="-"))
   return (a)
}

```

## API Access Functions

User friendly inputs for vet status:
`vetstatus = "all"` both non vet and vet business owners
`vetstatus = "vet"` vet business owners
`vetstatus = "none"`  non vet 


``` {r FxSwitchforCS}
#OLD - could not get if/switch to work then the whole thing failed

#GetVets<- function (vetstatus="*", sector="*") {

# if vetstatus = "all" ()
#   vetstatus <- "002&VET_GROUP=003&VET_GROUP=004"
# if vetstatus = "vet" {vetstatus <- "002"}
# if  vetstatus ="none" {vetstatus <- "004"}
       
  
  
# switch(vetstatus,
#     all = {vetstatus <- "&VET_GROUP=002&VET_GROUP=003&VET_GROUP=004"},
#     vet = {vetstatus <- "&VET_GROUP=002"},
#     none = {vetstatus <- "&VET_GROUP=004"}
#  #   stop ("Check input - Valid vetstatus are all, vet, none")
#     )  


```


# Research Questions 

## Company Summary Data
RQ1 -  Pull Vets from CS, all industries - 1 pull , Overview of VET owned Businesses by sector
Endpoint - Mod = Vet status

 Q2    - Pull from specific industry
 Endpoint mod = industry
These represents modification of 2 value on the endpoint

```{r CSDatavalidation echo=FALSE}
countpathcs<-"https://api.census.gov/data/2020/abscs?get=EMP,FIRMPDEMP,PAYANN&for=us:*&NAICS2017=*&key=324a30368aa24fca02eb737c15d7c32c63c891cb"
count_cs<-GET(countpathcs)
cnt_cs<-ConvertData(count_cs)
cntcs2<-CreateTibble(cnt_cs)
cntcs3<-cntcs2 %>% filter((nchar(NAICS2017)<3))
cntcs3
sum(as.double(cntcs3$FIRMPDEMP))
```



### RQ1 - Review of Vet Owned Businesses

Pull Data for all veterans to see where they have the largest presence across industry sectors.
```{r RQ1DataSun warnings=FALSE}
#This represent 1 modification of the CS endpoint, extracting by Vet status
#Works/Sat/Sun
#my_df <- GetVetsByInd(vetStatus = "002")
CS_AllVets<-CSData(vetStatus = "002", industryCode = "*")
#Return Status of API pull displayed

```

```{r RQ1SummariesSUN}
#sat/sun this works

#Total number of companies reporting vet owners:
#For Testing
sum(CS_AllVets$Firms)

#Create summarized dataset 
VetsBySector<-CS_AllVets %>%
  filter((substring(NAICS2017, 3,3)!="-")) %>% 
   group_by(SECTOR) %>% 
   summarize(Firms = sum(Firms))
#For Testing
  VetsBySector

#THis produces a proper graph by sector but dataset above is not summarized
#Plot of veterans by sector
g<- ggplot(VetsBySector,aes(SECTOR, Firms)) +
  geom_col()
g

```


#RQ2: Pull Data within the Transporation Industry

```{r RQ2DataSUN}
CS_48<-CSData(industryCode = "48*")
```

Analyze Transportation Sector:
```{r RQ2Summaries}

#Drop sector wide aggregations
CS_48a<-CS_48 %>% 
     filter((substring(NAICS2017, 3,3)!="-")) %>% 
     filter (VET_GROUP %in% c("002","003","004") )
       # mutate(VetStatus=as.factor(VET_GROUP), SubSector=as.factor(SUBSECTOR)) 
     
CS_SubSectors<-CS_48a %>%      
    group_by(SUBSECTOR,VET_GROUP) %>% 
     summarize(TotFirms = sum(Firms))
CS_SubSectors
 
#Works to show the largest subsectors employing vets and non vets alike
    
ggplot(CS_SubSectors, aes(SUBSECTOR,TotFirms, fill=VET_GROUP)) +
     geom_col()

CS_SubSectorEmps<-CS_48a %>%      
    group_by(SUBSECTOR,VET_GROUP) %>% 
     summarize(TotEmps = sum(Employees))
CS_SubSectorsEmps

ggplot(CS_SubSectorEmps, aes(SUBSECTOR,TotEmps, fill=VET_GROUP)) +
     geom_col()

```


A Cross Tabulation of the number of firms owned by Veterans vs non by Sector

```{r RQ2Crosstab}

#Because we have summarized data already, two options for showing the cross tab
#Un-summarize and display table (unnecessary)
#or
#Display the talbe using dplyr code
#Display  cross table

b<-CS_48 %>% 
 uncount(weights=.$Firms)
 b
table(b$SUBSECTOR,b$VET_GROUP)
 remove(b)
 

##Cross Tab using dplyr code
#The spread() 
a<-CS_48 %>% 
  # pivot_wider(names_from = VET_GROUP,values_from = Firms) %>% 
         group_by(SUBSECTOR,VET_GROUP) %>% 
         summarize(nFirms=sum(Firms)) %>% 
         #This is a superseded function, however it seems to work to efficiently create table 
         spread(VET_GROUP,nFirms,sep="_")
#Print cross tab table
a

```


Consider the pay rates within the transportation industry, are they different by vet vs. non vet owned firms?


```{r RQ2PayRates}
#WORKS 

pay_df<-CS_48 %>% 
  group_by(SUBSECTOR,VET_GROUP) %>% 
    summarize(totpayroll= sum(Payroll), TotEmps=sum(Employees)) %>% 
    mutate(AvgPayRate = totpayroll/TotEmps) %>% 
    filter(AvgPayRate!="NaN" & SUBSECTOR!="NA") %>% 
    filter(VET_GROUP %in% c("002","004"))
pay_df


ggplot(pay_df,aes(x=SUBSECTOR,y=AvgPayRate)) +
  #One Continuous value, shown by vet status 
   #  geom_density(alpha=.05,aes(fill=VET_GROUP), position="stack")
  # geom_boxplot(aes(color=VET_GROUP),fill="grey")
   geom_point(aes(colour=VET_GROUP))

ggplot(pay_df,aes(x=VET_GROUP,y=AvgPayRate)) +
  #One Continuous value, shown by vet status 
   #  geom_density(alpha=.05,aes(fill=VET_GROUP), position="stack")
   geom_boxplot(fill="grey")
 
ggplot(pay_df,aes(x=VET_GROUP,y=AvgPayRate)) +
  #One Continuous value, shown by vet status 
   #  geom_density(alpha=.05,aes(fill=VET_GROUP), position="stack")
   geom_boxplot(fill="grey")+
   facet_wrap(~SUBSECTOR)
```

--------------------
## Characteristics of Business Owners Data



```{r FXGetCBPath}

CBPath<- function(industryCode="*", vetStatus="*", bizAttrib="*", bizAnswer="*", sex="*") {
 #Wildcards used as defaults to pull these variables when not specified by user. 
  #Sunday verified
 
    endpath <- "https://api.census.gov/data/2020/abscb?"
    varlist <- "get=QDESC_LABEL,BUSCHAR_LABEL,EMP,EMPSZFI,EMPSZFI_LABEL,ETH_GROUP,ETH_GROUP_LABEL,FIRMPDEMP,PAYANN,RACE_GROUP,RACE_GROUP_LABEL,RCPPDEMP,SECTOR,SUBSECTOR,VET_GROUP_LABEL"

#star after param gets all records beginning with param value
  path <- paste0(endpath,varlist,
                "&for=us:*",
                 "&NAICS2017=",industryCode, 
                 "&VET_GROUP=", vetStatus, 
                 "&QDESC=",bizAttrib,
                 "&BUSCHAR=", bizAnswer,
                  "&SEX=", sex,
                 "&key=324a30368aa24fca02eb737c15d7c32c63c891cb")
 
   return(path)       
}        
```



```{r FxCBDatanew}
#Working
CBSun<- function (path) {

rawcb <- GET(path)
   
  #Print status code of returned data
   a<-status_code(rawcb)
   print(a)
  
#Tidy data
   my_cb<-rawcb %>%
     Wrangle() %>%
     ChangeVarsCB() %>% 
     DeAgto4()

   return(my_cb)
  }
```

### Research Questions from the  __Characteristics of Business__ (CB)endpoint  

The CB dataset allows us to better understand key attributes of business owners, for example whether businesses are family owned or franshcised. These can be accessed by selecting the QDESC survey question answers.

To refine this exploration, we can look at these answers sliced by industry, vet status, and ans sales receipts. This adds,3 additional modification points to the endpoint queries.  

Target mods to endpoint = 4

RQ2a&b: Get High level Industry data (by Sector?) and table for vets/not, vets/not by sex, race
RQ3: Get Vets by Industry, crosstabs on vets by industry, vet/sex by industry, vet/race by industry
RQ4: Pull on BO QDESC - FamBiz and Franchise
- RQ5:Pull by Vets, and Firmage to look at entrepreneurial efforts. - not a predicate var
Instead look at age with a pull using sales receipts under 500K


```{r CBDataValidation echo=FALSE}
#Data Validation
countpath<-"https://api.census.gov/data/2020/abscb?get=EMP,FIRMPDEMP,PAYANN&for=us:*&NAICS2017=*&key=324a30368aa24fca02eb737c15d7c32c63c891cb"
count_cb_df<-GET(countpath)
cnt_cb<-ConvertData(count_cb_df)
cnt2<-CreateTibble(cnt_cb)
cnt3<-cnt2 %>% filter((nchar(NAICS2017)>3) & (substring(NAICS2017, 3,3)!="-") )
cnt3
sum(as.double(cnt3$FIRMPDEMP))
```


Data pull notes: you can only pull for 1 slice at a time.
RQ_CB1  
__Family owned businesses to explore by sector - Mod = 1__


```{r}

familypath2<-CBPath(bizAttrib = "B02" )
fambiz<-CBSun(familypath2)



```

```{r}
fambizCNT<-fambiz %>% 
      filter(BUSCHAR %in% c("BM","BN"))    %>% 
      filter(NAICS2017 != "00") %>% 
     filter((substring(NAICS2017, 3,3)!="-")) %>% 
    filter((nchar(NAICS2017)>3)) %>% 
     select(SECTOR,SUBSECTOR, FIRMPDEMP,EMP,BUSCHAR, everything()) %>% 
    summarize(Firms = sum(FIRMPDEMP))
#Firm Count   
sum(fambizCNT$Firms)

famviz_cnt<-fambiz %>% 
      filter(BUSCHAR %in% c("BM","BN"))    %>% 
     select(SECTOR,SUBSECTOR, FIRMPDEMP,EMP,BUSCHAR, everything()) %>% 
     group_by(SECTOR) %>% 
    summarize(Firms = sum(FIRMPDEMP))
#Table
famviz_cnt
#By Sector and Sex
fam3<-fambiz %>% 
      filter(BUSCHAR %in% c("BM","BN"))    %>% 
     select(SECTOR,SUBSECTOR, FIRMPDEMP,EMP,BUSCHAR, everything()) %>% 
     group_by(SECTOR,SEX) %>% 
    summarize(Firms = sum(FIRMPDEMP))
#Table
fam3

#By Sector  Sex & Vet Status
#Sector and Sex
fam4<-fambiz %>% 
      filter(BUSCHAR %in% c("BM","BN") &
             (SEX %in% c("002","003","004")) &
             (VET_GROUP %in% c("002","003","004"))) %>% 
             
     select(SECTOR,SUBSECTOR, FIRMPDEMP,EMP,BUSCHAR, everything()) %>% 
     group_by(SECTOR,SEX,VET_GROUP) %>% 
    summarize(Firms = sum(FIRMPDEMP)) 
    #spread(VET_GROUP,Firms,sep="_")

#CONVERT THIS DF TO PLOT:
fam4


```


-----------------
__Family Business by Vet:__

Mod = 3
```{r warnings=FALSE}

vet_fambiz_path<-CBPath(vetStatus = "002",bizAttrib = "B02", bizAnswer = "BM")
vet_fambiz_df<-CBSun(vet_fambiz_path)
vf_df<-vet_fambiz_df %>% 
     select(SECTOR,SUBSECTOR, FIRMPDEMP,EMP,BUSCHAR, everything()) %>% 
    group_by(SECTOR) %>% 
   summarize(Firms = sum(FIRMPDEMP)) 
   # spread(BUSCHAR,Firms,sep="_")

#Table
vf_df
sum(vf_df$Firms)

```


Individual Facets:



__Family Business by Industry and Vet:__

```{r warnings=FALSE}
vet_fambiz_transp_path<-CBPath(vetStatus = "002",bizAttrib = "B02", bizAnswer = "BM", industryCode = "48*")
vet_fambiz_transp<-CBSun(vet_fambiz_transp_path)

vfbt<-vet_fambiz_transp %>% 
     select(SECTOR,SUBSECTOR, FIRMPDEMP,EMP,BUSCHAR, everything()) %>% 
    group_by(SECTOR) %>% 
   summarize(Firms = sum(FIRMPDEMP)) 
   # spread(BUSCHAR,Firms,sep="_")
vfbt
sum(vfbt$Firms)


```





Mod = 4
Analyze 
By Business Size within the transportation industry - belongs above in CS
```{r warnings=FALSE}
vet_fambiz_transp_path<-CBPath(vetStatus = "002",bizAttrib = "B02", bizAnswer = "BM", industryCode = "48*")
vet_fambiz_transp<-CBSun(vet_fambiz_transp_path)
```

```{r}
vt_df<-vet_fambiz_transp %>% 
     select(SECTOR,SUBSECTOR, FIRMPDEMP,EMP,BUSCHAR, everything()) %>% 
      group_by(SUBSECTOR) %>% 
   summarize(Firms = sum(FIRMPDEMP)) 
   # spread(BUSCHAR,Firms,sep="_")
vt_df
sum(vt_df$Firms)
```



__EXTRA__

```{r FamilyBiz}
familypath<-CB_FamilyPath(bizAttrib = "B02")
fb1<-GET(familypath)
fb2<-Wrangle(fb1)
fb3<-ChangeVarsCBFamily(fb2)
fb3
#Double Counts
sum(as.double(fb3$FIRMPDEMP))
```


```{r analyzeFamBiz}
f1_df<-fb3 %>% 
  filter(BUSCHAR %in% c("BM","BN"))    %>% 
  select(SECTOR,SUBSECTOR, FIRMPDEMP,EMP,BUSCHAR, everything()) %>% 
  group_by(SECTOR,BUSCHAR ) %>% 
   summarize(Firms = sum(FIRMPDEMP)) %>% 
    spread(BUSCHAR,Firms,sep="_")
f1_df

f1_count<-fb3 %>% 
      filter(BUSCHAR %in% c("BM","BN"))    %>% 
     select(SECTOR,SUBSECTOR, FIRMPDEMP,EMP,BUSCHAR, everything()) %>% 
    summarize(Firms = sum(FIRMPDEMP))
   
sum(f1_count$Firms)
#ADD IN PLOT

```




EXTRA IF 

```{r}
vetfambizpath<-CBVetFamPath(bizAttrib = "B02", vetStatus = "02")
fb1<-GET(familypath)
fb2<-Wrangle(fb1)
fb3<-ChangeVarsCBFamily(fb2)
fb3
```





```{r FxCB_FamilyPath}

CB_FamilyPath<- function(bizAttrib="*", bizAnswer="*") {
 #Wildcards used as defaults to pull these variables when not specified by user. 
  #Sunday verified
 
endpath <- "https://api.census.gov/data/2020/abscb?"
varlist <- "get=QDESC_LABEL,BUSCHAR_LABEL,EMP,FIRMPDEMP,PAYANN,RCPPDEMP,NAICS2017,SECTOR,SUBSECTOR"

#star after param gets all records beginning with param value
path <- paste0(endpath,varlist,
                "&for=us:*",
                 "&QDESC=",bizAttrib,
                 "&BUSCHAR=", bizAnswer,
                 "&key=324a30368aa24fca02eb737c15d7c32c63c891cb")
 
return(path)       
}        
```


```{r CBVetFamPath}

CBVetFamPath<- function(bizAttrib="*", bizAnswer="*", vetStatus="002") {
 #Wildcards used as defaults to pull these variables when not specified by user. 
  #Sunday verified
 
endpath <- "https://api.census.gov/data/2020/abscb?"
varlist <- "get=QDESC_LABEL,BUSCHAR_LABEL,EMP,FIRMPDEMP,PAYANN,RCPPDEMP,NAICS2017,SECTOR,SUBSECTOR"

#star after param gets all records beginning with param value
path <- paste0(endpath,varlist,
                "&for=us:*",
                 "&VET_GROUP=", vetStatus,
                 "&QDESC=",bizAttrib,
                 "&BUSCHAR=", bizAnswer,
                 "&key=324a30368aa24fca02eb737c15d7c32c63c891cb")
 
return(path)       
}        
```


```
