---
title: "temp"
author: "Lisa OConnell"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Introduction  
Data from the American Business Survey (ABS) conducted by the US CENSUS Bureau is explored in this vignette. The purpose of the vignette is to demonstrate access to multiple endpoints within the ABS data using a __X__ query parameters.  Additionally an Exploratory Data Analysis (EDA) is completed which includes both numeric and graphical summarizations.  This vignette was created and is shared using a GitHub repository and GitHub page.  


# Description of Data  
__Endpoints Over Time__  
The American Business Survey data represents aggregated responses from the ABS survey conducted in the prescribed year from the selected endpoing.  Aggregations are completed at several levels of data including>>>>

The ABS data accessed for this vignette represents the endpoints for the year 2021 which contains data from the 2020 survey.  Additional endpoints are available for the years 2020-2018.   


__Endpoint Pathways:__  
-  Company Summaries: http://api.census.gov/data/2020/abscs   
-  Characteristics of Business: https://api.census.gov/data/2020/abscb  
-  Characteristics of Business Owners: https://api.census.gov/data/2020/abscbo  
  
*Additional Endpoints Available*  
An additional endpoint is available, but is not used in this vignette:  

- Module Business Characteristics: https://api.census.gov/data/2020/absmcb  

__Data Format:__
As described in the ["CENSUS Data API User Guide"](https://www.census.gov/content/dam/Census/data/developers/api-user-guide/api-guide.pdf) The American Business Survey data returned through the API is a 2 dimensional JSON array formatted as shown below for 3 variables:  

[["EMP" ,"STATE", "PAYANN"]],  
["10000, "GA", "500000"],  
["5000, "TX", "650000"].... 


# Requirements: R Libraries  
The following libraries are used to process access and process data in R.  The `tidyverse` library provides a set of packages relevant for data tidying, wrangling, and summarization.  The `httr` package provides functions to work with urls.  The `jsonlite`package provides functions to work with json data.  

```{r libraries}
library(tidyverse)
library(httr)
library(jsonlite)
```


#Functions to Retrieve ABS Data

Several functions are created to facilitate data retrieval. These functions allow query arguements to be passed in to return specific results.  A wrapper function called `GetData` is created to contain the subfunctions needed to execute the API call (`GetAPIPath`,`GetMyRawData`), convert the data from JSON (INSERT FX), create a tidy tibble (`CreateTibble`), and reformat variables (`ChangeVars`).  

The `GetAPIPath` function allows for multiple arguements to be passed in to specify query terms.  For the two endpoints used in this vignette, the following query parameters are allowed:

*Characterics of Business data:*  
Query parameters:

__NAICS2017__   The NAICS industry code from the 2017 NAICS table.  Example industry sectors explored for this study are, XXXXX

__VET_GROUP__ A categorical variable that returns the status of the business owner as a veteran INSERT VALUES HERE

__SEX__ - A categorical variable returning the sex of the business owner.


```{r FXGetAPIPath}

GetAPIPathNAICS<- function(endpoint, param1){
 
if (!(endpoint %in% (c("CS","CB", "BO")))){ 
  stop("Check your input")}  

  #get Company Summary Data
  if (endpoint=="CS") { 
     endpath <- "https://api.census.gov/data/2020/abscs?"
      varlist <- "get=EMP,EMPSZFI,ETH_GROUP,FIRMPDEMP,
                 INDGROUP,INDLEVEL,PAYANN,RCPPDEMP,RCPSZFI,RACE_GROUP,
                 SECTOR,SEX,STATE,URSZFI,VET_GROUP,YEAR"
      #star after param gets all records beginning with param value
       path <- paste0(endpath,
                      varlist,
                      "&for=us:*&", 
                      param1, 
                      "*&key=324a30368aa24fca02eb737c15d7c32c63c891cb")
  }
  
#Get Characteristics of Business Data
   if (endpoint=="CB") { 
       endpath <- "https://api.census.gov/data/2020/abscb?"
       varlist <- "get=GEO_ID,NAME,NAICS2017_LABEL,BUSCHAR,
                   BUSCHAR_LABEL,CBSA,EMP,EMPSZFI,EMPSZFI_LABEL,
                   ETH_GROUP,ETH_GROUP_LABEL,FIRMPDEMP,INDGROUP,
                   INDLEVEL,PAYANN,QDESC,QDESC_LABEL,RACE_GROUP,
                  RACE_GROUP_LABEL,RCPPDEMP,RCPSZFI,RCPSZFI_LABEL,
                  SECTOR,SEX,STATE,SUBSECTOR,SUMLEVEL,URSZFI,
                  VET_GROUP,VET_GROUP_LABEL,YEAR,YIBSZFI,YIBSZFI_LABEL"
        #star after param gets all records beginnning with param value
        path <- paste0(endpath,varlist,
                       "&for=us:*&",
                       param1, 
                       "&key=324a30368aa24fca02eb737c15d7c32c63c891cb")
    }
  
#Get Business Owner Data  
  #parameter = vet status - var removed from below.
  #OWNER_VET=002 for vets
  if (endpoint=="BO") { 
      endpath<-"https://api.census.gov/data/2020/abscbo?"
      varlist <- "get=GEO_ID,NAME,NAICS2017,NAICS2017_LABEL,
                  OWNER_SEX,OWNER_SEX_LABEL,OWNER_ETH,OWNER_ETH_LABEL,
                  OWNER_RACE,OWNER_RACE_LABEL,OWNER_VET_LABEL,QDESC,
                  QDESC_LABEL,OWNCHAR,OWNCHAR_LABEL,YEAR,OWNPDEMP"
       path <- paste0(endpath,
                      varlist,
                      "&for=us:*&", 
                      param1,
                      "&key=324a30368aa24fca02eb737c15d7c32c63c891cb")
  }

return(path)
}

```


```{r FXgetrawdata}
#Working
#Gets the path to the endpoint for the query by NAICS
#Pulls the raw data
GetRawNAICS<-function (endpoint, param) {
    path<-GetAPIPathNAICS(endpoint, param)
    data<-GET(path)
    #Check status of returned data
    a<-status_code(data)
    #Code added - if statement
    #if (!(a==200)){
     #   print(a)
      #  stop("Data not Returned")
    #  }
    print(a)
    invisible(data)
}
```


```{r FxConvertData}
ConvertData<-function (myraw) {
  a11<-fromJSON(rawToChar(myraw$content))
  a111<-as_tibble(a11)
  return(a111)
}

```


```{r FxCreateTibble}
CreateTibble<-function(a_df) {
#Parse JSON data from API call
#q1<-content(my_rawdata,as="text", encoding="UTF-8")
#q11<-fromJSON(rawToChar(my_rawdata$content))
temp_df<-as_tibble(a_df)

#Get variable names from the first row of data
var_names<-slice(temp_df,1)
num_cols<-length(var_names)
b<-pivot_longer(var_names, cols=1:num_cols,values_to = "new_name", names_to="old_name")
#Replace variable names
my_names<-b$new_name
my_names

#Remove the column name header row
colnames(temp_df)<-my_names
new_data<-temp_df[-1, ]
invisible(new_data)

}
```


```{r FXChangeVars}
ChangeVarsCB<-function(my_df) {
#Converts all numeric variables from char to numeric
#Converts categorical variables to factors

#Create vector of numeric variables
number_vars<-c("PAYANN", "FIRMPDEMP", 
               "EMP", "RCPPDEMP")        
#Create vector of factor variables 
factor_vars<-c("EMPSZFI", "ETH_GROUP", "RCPSZFI",
               "URSZFI", "SEX", "VET_GROUP", "ETH_GROUP")
#Apply changes to my_df
my_df<-my_df %>% 
      mutate(across(factor_vars,as.factor)) %>% 
      mutate(across(number_vars, as.double))
return(my_df)
}  
```

```{r FXChangeVars}
ChangeVarsBO<-function(my_df) {
#Converts all numeric variables from char to numeric
#Converts categorical variables to factors
 
#Change the only numeric variable in the dataset
my_df$OWNPDEMP<-as.double(my_df$OWNPDEMP)
#Create vector of factor variables 
factor_vars<-c("OWNER_SEX", "OWNER_VET", 
               "OWNER_ETH","OWNER_RACE")
#Apply factor changes to my_df
my_df<-my_df %>% 
      mutate(across(factor_vars,as.factor))  
return(my_df)
}  
```

```{r FXChangeVars}
ChangeVars<-function(my_df,endpoint) {
#Converts all numeric variables from char to numeric
#Converts categorical variables to factors
  #COULD THIS BE CHAINED?
    if (endpoint%in% c("CS","CB")) { 
        #Create vector of numeric variables
        number_vars<-c("PAYANN", "FIRMPDEMP", 
                       "EMP", "RCPPDEMP")        
        #Create vector of factor variables 
        factor_vars<-c("EMPSZFI", "ETH_GROUP", "RCPSZFI",
                       "URSZFI", "SEX", "VET_GROUP", "ETH_GROUP")
        #Apply changes to my_df
        my_df<-my_df %>% 
          mutate(across(factor_vars,as.factor)) %>% 
          mutate(across(number_vars, as.double))
    }
  
  if (endpoint=="BO") { 
        #Change the only numeric variable in the dataset
        my_df$OWNPDEMP<-as.double(my_df$OWNPDEMP)
        #Create vector of factor variables 
        factor_vars<-c("OWNER_SEX", "OWNER_VET", 
                       "OWNER_ETH","OWNER_RACE")
        #Apply factor changes to my_df
        my_df<-my_df %>% 
            mutate(across(factor_vars,as.factor))  
  } 
   return(my_df)
}  
```


```{r datawrapperNAICS}
GetNAICS<-function(endpoint,param){
    a1<-GetRawNAICS(endpoint, param)
    a2<-ConvertData(a1)
    a3<-CreateTibble(a2)
    if (endpoint=="CB"){
        a4<ChangeVarsCB(a3)
       }
    if(endponit=="BO"){
        a4<-ChangeVarsBO(a3)
    }
    #a4<-ChangeVars(a3,endpoint)
#Return a dataset
 return(a4)    
}    

```
```{r}

```



To query only family owned business or non-family owned business, the following function is created.  This passes the desired type of firm into the function, and sets a default value for the industry of interest to the focus industry, transporation (NAICS code 484).  Family owned business are denoted by codes in survey questions(QDESC=B02) and responses to the B02 question = "BM" for Family Owned, "BN" for non family owned business.   

```{r}
#A Path generator function for family business queries
GetAPIPathFamBiz<- function(famcode){
 
#if (!(famcode %in% (c("BM","BN")))){ 
 # stop("Check your input")}  

#else{
  endpath <- "https://api.census.gov/data/2020/abscb?"
  varlist <- "get=GEO_ID,NAME,NAICS2017_LABEL,BUSCHAR_LABEL,CBSA,EMP,EMPSZFI,EMPSZFI_LABEL,ETH_GROUP,ETH_GROUP_LABEL,FIRMPDEMP,INDGROUP,INDLEVEL,PAYANN,QDESC_LABEL,RACE_GROUP,RACE_GROUP_LABEL,RCPPDEMP,RCPSZFI,RCPSZFI_LABEL,SECTOR,SEX,STATE,SUBSECTOR,SUMLEVEL,URSZFI,VET_GROUP,VET_GROUP_LABEL,YEAR,YIBSZFI,YIBSZFI_LABEL"
  #star after param gets all records beginnning with param value
   path <- paste0(endpath,varlist,
                       "&for=us:*&QDESC=B02",
                       "&BUSCHAR=",
                       famcode,
                       "&NAICS2017=484",
                       "&key=324a30368aa24fca02eb737c15d7c32c63c891cb")
#    }

return(path)
}

```


```{r datawrapperFamilyOwned}
#Calls the path generator function, then wrangles data.
#This call is only applicable to the  Characteristics of Businesses Endpoint
GetFamBiz<-function(famcode){
    a1<-GetRawFamBiz(famcode)
    a2<-ConvertData(a1)
    a3<-CreateTibble(a2)
    a4<-ChangeVarsCB(a3)
#Return a dataset
 return(a4)   
}
```



```{r}
#Working
#Gets the path to the endpoint for the query by NAICS
#Pulls the raw data
GetRawFamBiz<-function (famcode) {
    path<-GetAPIPathFamBiz(famcode)
    data<-GET(path)
    #Check status of returned data
    a<-status_code(data)
    #Code added - if statement
    #if (!(a==200)){
     #   print(a)
      #  stop("Data not Returned")
    #  }
    print(a)
    invisible(data)
}
```

```{rvetpatCB}

```


```{r}
GetVetsRawCB<-function(vetcode, naics="&NAICS2017=484"){
#Create URL path
endpath <- "https://api.census.gov/data/2020/abscb?"
varlist <- "get=GEO_ID,NAME,NAICS2017_LABEL,BUSCHAR_LABEL,CBSA,EMP,EMPSZFI,EMPSZFI_LABEL,ETH_GROUP,ETH_GROUP_LABEL,FIRMPDEMP,INDGROUP,INDLEVEL,PAYANN,QDESC_LABEL,RACE_GROUP,RACE_GROUP_LABEL,RCPPDEMP,RCPSZFI,RCPSZFI_LABEL,SECTOR,SEX,STATE,SUBSECTOR,SUMLEVEL,URSZFI,VET_GROUP_LABEL,YEAR,YIBSZFI,YIBSZFI_LABEL"
#star after param gets all records beginnning with param value
path <- paste0(endpath,varlist,
                      "&for=us:*&", 
                      "&VET_GROUP=" ,
                       vetcode,
                       "&NAICS2017=484",
                       "&key=324a30368aa24fca02eb737c15d7c32c63c891cb")
data<-GET(path)
#Check status of returned data
a<-status_code(data)
#Code added - if statement
    #if (!(a==200)){
     #   print(a)
      #  stop("Data not Returned")
    #  }
print(a)
invisible(data)
}
```



```{r datawrapperFamilyOwned}
#Calls the path generator function, then wrangles data.
#This call is only applicable to the  Characteristics of Businesses Endpoint
GetVets<-function(vetcode,naics){
    a1<-GetVetsRawCB(vetcode,naics)
    a2<-ConvertData(a1)
    a3<-CreateTibble(a2)
    a4<-ChangeVarsCB(a3)
#Return a dataset
 return(a4)   
}
```


BY SEX:

```{r}
#A Path generator function for family business queries
GetAPIPathSx<- function(codevalue){
 
#if (!(famcode %in% (c("BM","BN")))){ 
 # stop("Check your input")}  

#else{
  endpath <- "https://api.census.gov/data/2020/abscb?"
  varlist <- "get=GEO_ID,NAME,NAICS2017_LABEL,BUSCHAR_LABEL,CBSA,EMP,EMPSZFI,EMPSZFI_LABEL,ETH_GROUP,ETH_GROUP_LABEL,FIRMPDEMP,INDGROUP,INDLEVEL,PAYANN,QDESC_LABEL,RACE_GROUP,RACE_GROUP_LABEL,RCPPDEMP,RCPSZFI,RCPSZFI_LABEL,SECTOR,STATE,SUBSECTOR,SUMLEVEL,URSZFI,VET_GROUP,VET_GROUP_LABEL,YEAR,YIBSZFI,YIBSZFI_LABEL"
  #star after param gets all records beginnning with param value
   path <- paste0(endpath,varlist,
                       "&for=us:*&", 
                       "SEX=",
                       codevalue,
                       "&NAICS2017=484",
                       "&key=324a30368aa24fca02eb737c15d7c32c63c891cb")
#    }

return(path)
}

```


```{r}
#Working
#Gets the path to the endpoint for the query by NAICS
#Pulls the raw data
GetRawBySx<-function (codein) {
    path<-GetAPIPathSx(codein)
    data<-GET(path)
    #Check status of returned data
    a<-status_code(data)
    #Code added - if statement
    #if (!(a==200)){
     #   print(a)
      #  stop("Data not Returned")
    #  }
    print(a)
    invisible(data)
}
```

```{r datawrapperFamilyOwned}
#Calls the path generator function, then wrangles data.
#This call is only applicable to the  Characteristics of Businesses Endpoint
GetbySx<-function(codein){
    a1<-GetRawBySx(codein)
    a2<-ConvertData(a1)
    a3<-CreateTibble(a2)
    a4<-ChangeVarsCB(a3)
#Return a dataset
 return(a4)   
}
```


-------

Access the Company Summary Endpoint 
```{r GetCSData}
#REMOVE - DEMONSTRATION OF CS DATA

CS_NEW2<-GET("https://api.census.gov/data/2020/abscs?get=EMP,EMPSZFI,ETH_GROUP,FIRMPDEMP,INDGROUP,INDLEVEL,PAYANN,RCPPDEMP,RCPSZFI,RACE_GROUP,SECTOR,SEX,STATE,URSZFI,VET_GROUP,YEAR&for=us:*&NAICS2017=71*&key=324a30368aa24fca02eb737c15d7c32c63c891cb")
#Check request successful via code
status_code(CS_NEW2)
#Get data to parse, 
#           q1<-content(CSQ,as="parsed", encoding="UTF-8")
cs11<-fromJSON(rawToChar(CS_NEW2$content))
cs111<-as_tibble(cs11)

#Get variable names from the first row of data
var_names<-slice(cs111,1)
num_cols<-length(var_names)
b<-pivot_longer(var_names, cols=1:num_cols,values_to = "new_name", names_to="old_name")

#Replace variable names
my_names<-b$new_name
my_names
colnames(cs111)<-my_names

#Remove the column name header row from the data
cs1111<-cs111[-1, ]



```


Access the Characteristics of Business Endpoint 

```{r CBdata}

#REMOVE - DEMONSTRATION OF CB DATA
#char_biz_path<-c("https://api.census.gov/data/2020/abscb")

#char_biz_test<-GET("https://api.census.gov/data/2020/abscb?get=GEO_ID,NAME,NAICS2017,NAICS2017_LABEL,SEX,SEX_LABEL,ETH_GROUP,ETH_GROUP_LABEL,RACE_GROUP,RACE_GROUP_LABEL,VET_GROUP,VET_GROUP_LABEL,QDESC,QDESC_LABEL,BUSCHAR,BUSCHAR_LABEL,YEAR,FIRMPDEMP,FIRMPDEMP_F,FIRMPDEMP_PCT,FIRMPDEMP_PCT_F,RCPPDEMP,RCPPDEMP_F,RCPPDEMP_PCT,RCPPDEMP_PCT_F,EMP,EMP_F,EMP_PCT,EMP_PCT_F,PAYANN,PAYANN_F,PAYANN_PCT,PAYANN_PCT_F,FIRMPDEMP_S,FIRMPDEMP_S_F,FIRMPDEMP_PCT_S,FIRMPDEMP_PCT_S_F,RCPPDEMP_S,RCPPDEMP_S_F,RCPPDEMP_PCT_S,RCPPDEMP_PCT_S_F,EMP_S,EMP_S_F,EMP_PCT_S,EMP_PCT_S_F,PAYANN_S,PAYANN_S_F,PAYANN_PCT_S,PAYANN_PCT_S_F&for=us:*&QDESC_LABEL=SPOUSES&key=324a30368aa24fca02eb737c15d7c32c63c891cb")
#status_code(char_biz_test)

#WORKING CODE
CBQt<-GET("https://api.census.gov/data/2020/abscb?get=BUSCHAR,CBSA,EMP_PCT,EMPSZFI,ETH_GROUP,FIRMPDEMP,INDGROUP,INDLEVEL,PAYANN,QDESC,RACE_GROUP,RCPPDEMP,RCPSZFI,SECTOR,SEX,STATE,SUBSECTOR,SUMLEVEL,URSZFI,VET_GROUP,YEAR,YIBSZFI&for=us:*&NAICS2017=71&key=324a30368aa24fca02eb737c15d7c32c63c891cb")

status_code(CBQt)
#Get data to parse, 
b1<-content(CBQt,as="parsed", encoding="UTF-8")
b11<-fromJSON(rawToChar(CBQt$content))
b111<-as_tibble(q11)

#Get variable names from the first row of data
var_names<-slice(b111,1)
num_cols<-length(var_names)
b<-pivot_longer(var_names, cols=1:num_cols,values_to = "new_name", names_to="old_name")
#Replace variable names
my_names<-b$new_name
my_names


#dimnames(my_ChB)[[2]]<-my_names
colnames(b111)<-my_names
#Remove the column name header row
b1111<-b111[-1, ]

```

Access the Business Owners Endpoint
```{r}
#REMOVE - DEMONSTRATION OF BO DATA
#WORKING CODE EXAMPLE
#WORKING CODE EXAMPLE PARAM: OWNER_VET=002, 

cbo_test<-GET("https://api.census.gov/data/2020/abscbo?get=GEO_ID,NAME,NAICS2017,NAICS2017_LABEL,OWNER_SEX,OWNER_SEX_LABEL,OWNER_ETH,OWNER_ETH_LABEL,OWNER_RACE,OWNER_RACE_LABEL,OWNER_VET,OWNER_VET_LABEL,QDESC,QDESC_LABEL,OWNCHAR,OWNCHAR_LABEL,YEAR,OWNPDEMP,OWNPDEMP_F,OWNPDEMP_PCT,OWNPDEMP_PCT_F,OWNPDEMP_S,OWNPDEMP_S_F,OWNPDEMP_PCT_S,OWNPDEMP_PCT_S_F&for=us:*&QDESC_LABEL=YRACQBUS&key=324a30368aa24fca02eb737c15d7c32c63c891cb")
status_code(CBO_test)


CBOvet3<-GET("https://api.census.gov/data/2020/abscbo?get=GEO_ID,NAME,NAICS2017,NAICS2017_LABEL,OWNER_SEX,OWNER_SEX_LABEL,OWNER_ETH,OWNER_ETH_LABEL,OWNER_RACE,OWNER_RACE_LABEL,OWNER_VET,OWNER_VET_LABEL,QDESC,QDESC_LABEL,OWNCHAR,OWNCHAR_LABEL,YEAR,OWNPDEMP&for=us:*&OWNER_VET=002&key=324a30368aa24fca02eb737c15d7c32c63c891cb")
status_code(CBOvet3)

#Get data to parse, 
o1<-content(CBOvet,as="parsed", encoding="UTF-8")
o11<-fromJSON(rawToChar(CBOvet$content))
o111<-as_tibble(o11)

#Get variable names from the first row of data
var_names<-slice(o111,1)
num_cols<-length(var_names)
b<-pivot_longer(var_names, cols=1:num_cols,values_to = "new_name", names_to="old_name")
#Replace variable names
my_names<-b$new_name
my_names

colnames(o111)<-my_names
#Remove the column name header row
o1111<-o111[-1, ]


```


Modifications to the first endpoint - Characteristics of Businesses Surveyed  
- Select the industry Code to investigated  
- Sales receipts over a certain level  

 
# Data Exploration  

## Data Slices from the  __Characteristics of Business__ (CB)endpoint  

For this vignette, I am interested in exploring data representing the truck transportation industry.  I am interested in undertanding a general overview, and a focus on veteran owned businesses in particular.   

This industry is represented by North American Indsutry Classification System ([NAICS](https://www.census.gov/naics/?99967)) codes beginning with 484 - Truck Transportation. Records are stored at a lower level of classification as well as the aggregate level - 484.  

To create an overview of the truck transportation industry, data is extracted from the CB endpoint.  

```{r}
#Use wrapper function to pull data from the CB endpoint with NAICS_2017 query.  
#The "*" following 484 is a wildcard that will pull all records that start with 484.  
TRUCKS<-GetNAICS("CB","NAICS_2017=484*")
```

__Data Slices From Characteristics of Business Owners Endpoint__
To get an overview of business owners in general and veteran business owners in particular, I have queried the Characteristics of Business Owners endpoint for the transportation industry.  This query is refined further to 

```{r}
FamBiz<-GetFamBiz(famcode = "BM")
```

Veteran Owned Business

```{r}
Vets<-GetVets(vetcode = "002")

  
  
```

By Sex

```{r}
Females<-GetbySx(codein = "002")
```




