---
title: "temp"
author: "Lisa OConnell"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description of Data
2020 business survey
2017 naics codes

__Endpoint Pathways:__  
-  Company Summaries: http://api.census.gov/data/2020/abscs   
  -  Characteristics of Business: https://api.census.gov/data/2020/abscb  
-  Characteristics of Business Owners: https://api.census.gov/data/2020/abscbo  
*Additional Endpoints Available*  
An additional endpoint is available, but is not used in this vignette:  

- Module Business Characteristics: https://api.census.gov/data/2020/absmcb  

# R Libraries
The following libraries are used to process access and process data in R.  The `tidyverse` library provides a set of packages relevant for data tidying, wrangling, and summarization.  The `httr` package provides functions to work with urls.  The `jsonlite`package provides functions to work with json data.  

```{r}
library(tidyverse)
library(httr)
library(jsonlite)
```


#Function to Retrieve data

```{r FXgetdatapath}

GetAPICallbyNC<- function(endpoint, param){
 
if (!(endpoint %in% (c("CS","CB", "BO")))){ 
  stop("Check your input")}  
#get Company Summary Data
  if (endpoint=="CS") { 
     endpath <- "https://api.census.gov/data/2020/abscs?"
  
      varlist <- "get=EMP,EMPSZFI,ETH_GROUP,FIRMPDEMP,INDGROUP,INDLEVEL,PAYANN,RCPPDEMP,RCPSZFI,RACE_GROUP,SECTOR,SEX,STATE,URSZFI,VET_GROUP,YEAR"
      #star after param gets all records beginnning with param value
       path <- paste0(endpath,varlist,"&for=us:*&", param, "*&key=324a30368aa24fca02eb737c15d7c32c63c891cb")
  }
  
#Get Characteristics of Business Data
   if (endpoint=="CB") { 
       endpath <- "https://api.census.gov/data/2020/abscb?"
       varlist <- "get=BUSCHAR,CBSA,EMP_PCT,EMPSZFI,ETH_GROUP,FIRMPDEMP,INDGROUP,INDLEVEL,PAYANN,QDESC,RACE_GROUP,RCPPDEMP,RCPSZFI,SECTOR,SEX,STATE,SUBSECTOR,SUMLEVEL,URSZFI,VET_GROUP,YEAR,YIBSZFI"
        #star after param gets all records beginnning with param value
        path <- paste0(endpath,varlist,"&for=us:*&", param, "*&key=324a30368aa24fca02eb737c15d7c32c63c891cb")
    }
  
#Get Business Owner Data  
  #parameter = vet status 
  #OWNER_VET=002 for vets
  if (endpoint=="BO") { 
   endpath<-"https://api.census.gov/data/2020/abscbo?"
   varlist <- "get=GEO_ID,NAME,NAICS2017,NAICS2017_LABEL,OWNER_SEX,OWNER_SEX_LABEL,OWNER_ETH,OWNER_ETH_LABEL,OWNER_RACE,OWNER_RACE_LABEL,OWNER_VET,OWNER_VET_LABEL,QDESC,QDESC_LABEL,OWNCHAR,OWNCHAR_LABEL,YEAR,OWNPDEMP"
    path <- paste0(endpath,varlist,"&for=us:*&", param, "&key=324a30368aa24fca02eb737c15d7c32c63c891cb")
  }

return(path)
}

```



```{r FXgetrawdata}
#Working
GetMyRawData<-function (endpoint, param) {
    path<-GetAPICallbyNC(endpoint, param)
    #DELANEY - 3 THIS FAILS AND I THINK IT MAY BE BECAUSE OF HOW I AM PASSING STRING ...do the "" fall off?
    data<-GET(path)
    
    a<-status_code(data)
    print(a)
    invisible(data)
    #return(data)
}

```



```{r}
ConvertData<-function (myraw) {
  a11<-fromJSON(rawToChar(myraw$content))
  a111<-as_tibble(a11)
}

```


```{r parsedataTOTibble}
CreateTibble<-function(a_df) {
#Parse JSON data from API call
#q1<-content(my_rawdata,as="text", encoding="UTF-8")
#q11<-fromJSON(rawToChar(my_rawdata$content))
temp_df<-as_tibble(a_df)

#Get variable names from the first row of data
var_names<-slice(temp_df,1)
num_cols<-length(var_names)
b<-pivot_longer(var_names, cols=1:num_cols,values_to = "new_name", names_to="old_name")
#Replace variable names
my_names<-b$new_name
my_names

#Remove the column name header row
colnames(temp_df)<-my_names
new_data<-temp_df[-1, ]
invisible(new_data)

}
```

```{r datawrapper}
GetData<-function(endpoint,param){
    a1<-GetMyRawData(endpoint, param)
    a2<-ConvertData(a1)
    a3<-CreateTibble(a2)
#Return a dataset
 return(a3)    
}    

```


```{r GetCSData}
#DELANEY 1- THIS WORKS!
#WORKING CODE FO CS ENDPOINTS
#Parm with Data, NCAIS2017=71


CS_NEW2<-GET("https://api.census.gov/data/2020/abscs?get=EMP,EMPSZFI,ETH_GROUP,FIRMPDEMP,INDGROUP,INDLEVEL,PAYANN,RCPPDEMP,RCPSZFI,RACE_GROUP,SECTOR,SEX,STATE,URSZFI,VET_GROUP,YEAR&for=us:*&NAICS2017=71*&key=324a30368aa24fca02eb737c15d7c32c63c891cb")
#Check request successful via code
status_code(CS_NEW2)
#Get data to parse, 
#           q1<-content(CSQ,as="parsed", encoding="UTF-8")
cs11<-fromJSON(rawToChar(CS_NEW2$content))
cs111<-as_tibble(cs11)

#Get variable names from the first row of data
var_names<-slice(cs111,1)
num_cols<-length(var_names)
b<-pivot_longer(var_names, cols=1:num_cols,values_to = "new_name", names_to="old_name")

#Replace variable names
my_names<-b$new_name
my_names
colnames(cs111)<-my_names

#Remove the column name header row from the data
cs1111<-cs111[-1, ]



```



Access the Characteristics of Business Endpoint 

```{r CBdata}
#char_biz_path<-c("https://api.census.gov/data/2020/abscb")

#char_biz_test<-GET("https://api.census.gov/data/2020/abscb?get=GEO_ID,NAME,NAICS2017,NAICS2017_LABEL,SEX,SEX_LABEL,ETH_GROUP,ETH_GROUP_LABEL,RACE_GROUP,RACE_GROUP_LABEL,VET_GROUP,VET_GROUP_LABEL,QDESC,QDESC_LABEL,BUSCHAR,BUSCHAR_LABEL,YEAR,FIRMPDEMP,FIRMPDEMP_F,FIRMPDEMP_PCT,FIRMPDEMP_PCT_F,RCPPDEMP,RCPPDEMP_F,RCPPDEMP_PCT,RCPPDEMP_PCT_F,EMP,EMP_F,EMP_PCT,EMP_PCT_F,PAYANN,PAYANN_F,PAYANN_PCT,PAYANN_PCT_F,FIRMPDEMP_S,FIRMPDEMP_S_F,FIRMPDEMP_PCT_S,FIRMPDEMP_PCT_S_F,RCPPDEMP_S,RCPPDEMP_S_F,RCPPDEMP_PCT_S,RCPPDEMP_PCT_S_F,EMP_S,EMP_S_F,EMP_PCT_S,EMP_PCT_S_F,PAYANN_S,PAYANN_S_F,PAYANN_PCT_S,PAYANN_PCT_S_F&for=us:*&QDESC_LABEL=SPOUSES&key=324a30368aa24fca02eb737c15d7c32c63c891cb")
#status_code(char_biz_test)

#WORKING CODE
CBQt<-GET("https://api.census.gov/data/2020/abscb?get=BUSCHAR,CBSA,EMP_PCT,EMPSZFI,ETH_GROUP,FIRMPDEMP,INDGROUP,INDLEVEL,PAYANN,QDESC,RACE_GROUP,RCPPDEMP,RCPSZFI,SECTOR,SEX,STATE,SUBSECTOR,SUMLEVEL,URSZFI,VET_GROUP,YEAR,YIBSZFI&for=us:*&NAICS2017=71&key=324a30368aa24fca02eb737c15d7c32c63c891cb")

status_code(CBQt)
#Get data to parse, 
b1<-content(CBQt,as="parsed", encoding="UTF-8")
b11<-fromJSON(rawToChar(CBQt$content))
b111<-as_tibble(q11)

#Get variable names from the first row of data
var_names<-slice(b111,1)
num_cols<-length(var_names)
b<-pivot_longer(var_names, cols=1:num_cols,values_to = "new_name", names_to="old_name")
#Replace variable names
my_names<-b$new_name
my_names


#dimnames(my_ChB)[[2]]<-my_names
colnames(b111)<-my_names
#Remove the column name header row
b1111<-b111[-1, ]

```

Access the Business Owners Endpoint
```{r}

#WORKING CODE EXAMPLE
#WORKING CODE EXAMPLE PARAM: OWNER_VET=002, 

cbo_test<-GET("https://api.census.gov/data/2020/abscbo?get=GEO_ID,NAME,NAICS2017,NAICS2017_LABEL,OWNER_SEX,OWNER_SEX_LABEL,OWNER_ETH,OWNER_ETH_LABEL,OWNER_RACE,OWNER_RACE_LABEL,OWNER_VET,OWNER_VET_LABEL,QDESC,QDESC_LABEL,OWNCHAR,OWNCHAR_LABEL,YEAR,OWNPDEMP,OWNPDEMP_F,OWNPDEMP_PCT,OWNPDEMP_PCT_F,OWNPDEMP_S,OWNPDEMP_S_F,OWNPDEMP_PCT_S,OWNPDEMP_PCT_S_F&for=us:*&QDESC_LABEL=YRACQBUS&key=324a30368aa24fca02eb737c15d7c32c63c891cb")
status_code(CBO_test)


CBOvet3<-GET("https://api.census.gov/data/2020/abscbo?get=GEO_ID,NAME,NAICS2017,NAICS2017_LABEL,OWNER_SEX,OWNER_SEX_LABEL,OWNER_ETH,OWNER_ETH_LABEL,OWNER_RACE,OWNER_RACE_LABEL,OWNER_VET,OWNER_VET_LABEL,QDESC,QDESC_LABEL,OWNCHAR,OWNCHAR_LABEL,YEAR,OWNPDEMP&for=us:*&OWNER_VET=002&key=324a30368aa24fca02eb737c15d7c32c63c891cb")
status_code(CBOvet3)

#Get data to parse, 
o1<-content(CBOvet,as="parsed", encoding="UTF-8")
o11<-fromJSON(rawToChar(CBOvet$content))
o111<-as_tibble(o11)

#Get variable names from the first row of data
var_names<-slice(o111,1)
num_cols<-length(var_names)
b<-pivot_longer(var_names, cols=1:num_cols,values_to = "new_name", names_to="old_name")
#Replace variable names
my_names<-b$new_name
my_names

colnames(o111)<-my_names
#Remove the column name header row
o1111<-o111[-1, ]


```



Function to Change variable types  
- There are 4 interger variables of interest in the Company Summaries endpoint:  
 *EMP*: Number of employees  
 *FIRMPDEMP*: The number of employer firms  
 *PAYANN*: The annual payroll  
 *RCPPDEMP*: Sales, shipments or revenue in ($000s)  
 

```{r FXconverttonumeric}
#WORKING
CV<-function(my_df,endpoint) {
#Check User Input for valid values.
  #Converts all numeric variables from char to numeric
  #working as of 10/4
if (!(endpoint %in% c("CS","CB", "BO"))){ stop("Check your input")}  
if (endpoint=="CS") { 
   my_df$PAYANN<-as.double(my_df$PAYANN)
   my_df$FIRMPDEMP<-as.double(my_df$FIRMPDEMP)
   my_df$EMP<-as.double(my_df$EMP)
   my_df$RCPPDEMP<-as.double(my_df$RCPPDEMP)
    }
if (endpoint=="CB") { 
   my_df$PAYANN<-as.double(my_df$PAYANN)
   my_df$FIRMPDEMP<-as.double(my_df$FIRMPDEMP)
   my_df$EMP<-as.double(my_df$EMP)
   my_df$RCPPDEMP<-as.double(my_df$RCPPDEMP)
    }
  if (endpoint=="BO") { 
   my_df$OWNPDEMP<-as.double(my_df$OWNPDEMP)
       } 
  
   return(invisible(my_df))
}  
  
```

